---
title: "Initial Korean covid 19"
author: "Ashley Cook"
date: "June 21, 2023"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
kRdata <- kRdata %>%
mutate(across(where(is.character), as.factor))

```
### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 2020
### This data was retrieved from a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
  
# Your Question - Is dying independent of sex?

## METHODS
  
We will analyze the sex of the patients and the result if they are alive or dead. I will use the variables in my collection of Cat~cat. This is a type of problem that we can use the helper file and we will copy into our project to get the correct answer. 



## Graphical results - use ggplot commands to do this  

```{r}
dd2 <- kRdata %>% group_by(result,Sex) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=result,y=count,fill=Sex))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=result,y=prcnt*100,fill=Sex)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

  
There is more females than males that show on these charts. Twice the amount of people that are males that are dead as there is females, which is pretty odd since there are more females than males. It is killing males on a 2 to 1 basis. In our graphical results, Covid wants to kill more of the males than females. 

  
## Numerical results

```{r}
table1 <- xtabs(~result + Sex, data=kRdata)
rowPerc(table1)
colPerc(table1)
```
  
There is a total of 100 men and women each. The females that are alive are 56% of them, while males are 44% alive. Then the males that are dead is a percentage of 64, while females have a percentage of 36. In the second chart, below the first, shows that there is a percentage of females that is 98.7, while men have a percentage of 97. The amount of females that are dead are 1.4, while males have 3.

## Inferential results 

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

  
In the group of 1831females were alive and 1411males. All in all more females are alive than male. The p value is 1 out of a thousand. It is a small chance to have females dead over this disease than males. In fisher's test 9 out of 10,000 that are about the same p value. You can also see the odds ratio of. The odds of female living against the odds of males dying. We want to know the odds of this happening. 

## Conclusions
  
Overall, there was a lot more females then there were males, but males having the higher death percentages from Covid than females. Till this day we still try to figure out why males were dying from the disease over females, is it because they got out more than females? The plots that we created showed us the results of the amount of females and males that died from Covid. The final result is that we want to know what the odds are of this happening. Even with the males having a lower percentage of them being alive than females, they still had the highest percentage of deceased.
